Efficient Handling of String-Number Conversion

### 2020-03-30 final-version/evaluation.tex
little changes in the text

We have implemented our string constraint solving procedure in a tool called {\tool}. {\tool} is implemented as a theory solver of the SMT solver Z3~\cite{z3}. In this way, we can concentrate on solving conjunctive constraints and let Z3 handle the other boolean connectives. Secondly, it makes it possible to solve not only formulae over string constraints but also combinations of string constraints with other theories that Z3 supports. Furthermore, this approach allows us to more effectively handle the arithmetic constraints that are generated by the under-approximation module and eliminate the need to have our own parser for input formulae.

In {\tool}, we use the following PFA selection strategy. First, we use \emph{numeric PFAs} for string variables appearing in string-number conversion and \emph{standard PFAs} for others. Then we select size $m$ for numeric PFAs and $p$ loops of size $q$ for standard PFAs. Initially, we set $(m,p,q)=(5,2,q)$, where $q$ is dynamic and achieved from our internal static analysis. We double $m$, increase $p$ and $q$ by one if more precise refinement is required. We set an upper bound for each parameter and report UNKNOWN if a solution cannot be found within the bound.

Our over-approximation module uses also some heuristics to derive the constant value of any side of the constraint $n=\sti{x}$ to refine the over-approximation. For instance, assume in some case, we could derive that the value of $n=12$ from some integer constraints.  Then, we could further derive the value of $x$ is in the regular language $(0^*12)$.

The way our theory solver and Z3 interacts is almost standard. Every time when Z3 asks our theory solver a string constraint satisfiability problem, our solver tries to prove it is SAT or UNSAT using the aforementioned procedure. For under-approximation, every time when a corresponding linear formula is created, we attach the current value of $m$, $p$, $q$ with the formula, and then push it to Z3 core. If our theory solver reports UNKNOWN, Z3 remembers it in a global flag \textsf{incomplete} and either tries another solution branch, or the same solution branch with different value of $m$, $p$, $q$. If Z3 completes the search of all solution branches, it reports UNSAT if the flag \textsf{incomplete} is down, and UNKNOWN otherwise.

@@

We perform two sets of experiments. In the first set of experiments, we compare {\tool} with other tools on existing benchmarks over basic string constraints. Those benchmarks do not involve string-number conversion function. In the second set of experiments, we compare {\tool} with other tools on new suites focusing on string-number conversion. Our goals of experiments are the following:
\smallskip

\begin{itemize}
	\item {\tool} performs as good as, or better than other tools in solving the  satisfiability problems of basic string constraints.

@@

	\item {\tool} performs significantly better than other tools in solving the satisfiability problems on string-number conversion benchmark, and this shows  the efficiency of PFA in general and \emph{numeric} PFA in particular.
\end{itemize}
		\smallskip

@@

\begin{itemize}
	\item PyEx~\cite{pyex} is from running the symbolic executor PyEx over some Python packages.

@@

	\item \changed{ LeetCode is from running PyEx over some sample code collected from the LeetCode~\cite{LeetCode} website, including functions that check if a string is a valid IPv4 or IPv6 address, sum up two binary numbers, check if an input string is an abbreviation of another input string, and decode a sequence of digits to strings according to a given mapping.}

@@

	\item StringFuzz~\cite{blotsky2018stringfuzz}  is generated from the fuzz testing tool of the same name.

@@

In the second experiment, we compare with tools supporting string-number conversion on the benchmarks collected from the symbolic executor \texttt{Py-Conbyte}\footnote{\url{https://github.com/spencerwuwu/py-conbyte}}, which has the supports of string-number conversion. We ran it on several examples collected from the LeetCode platform and from Python core libraries, which involve diverse usages of string-number conversion in Python such as parsing date-time, verifying and restoring IP addresses from strings, etc. We also have examples that encode execution paths of some JavaScript programs (the Luhn algorithm and some array manipulation).

All experiments were executed on a machine with 4-core CPU, 8 GiB RAM. The timeout was set to 10s for each test.
We use the results from {\tool}, CVC4, and Z3 as the reference answer for the validation of the correctness of the results. Occasionally, two of them report inconsistent (one SAT and one UNSAT) answers to the same test. To decide which solver reports the incorrect answer, we developed a validator, which takes the model $I$ returned from the solver who reported SAT, assigns $I(x)$ to all variables $x$ in the test to obtain a modified test, and re-evaluates the modified test by multiple solvers. If the results from all solvers are consistent, we mark the test SAT or UNSAT according to the results. Otherwise, we manually simplify and inspect the test until we get a conclusive result.

The results of the experiments are summarized in Table~\ref{table:base_benchmark}, Table~\ref{table:str_int_benchmark}, and Table~\ref{table:checkLuhn}. Rows with heading \texttt{SAT}/\texttt{UNSAT} show numbers of solved formulae. Rows with heading \texttt{UNKNOWN}, \texttt{TIMEOUT}, or \texttt{ERROR} indicate the number of instances the solver fails to return an answer. \texttt{INCORRECT} shows the number of cases that a tool gives wrong answer\footnote{A correct answer is decided by mainly CVC4, then the consensus of {\tool} or Z3.}.

@@

From Table~\ref{table:base_benchmark}, we can see that the performance of {\tool} is as good as the most competitive tools such as CVC4 and Z3 on basic string constraints. In all of the benchmarks, {\tool} ranked either the 1st or the 2nd on the number of solved (SAT+UNSAT) cases. If we look into scrutiny, for the APLAS and StringFuzz benchmarks, there is a significant difference between {\tool} and the top-ranked tool on the SAT cases. But we believe this is not very important because both these two benchmarks are artificial, either hand-crafted or randomly generated. The more critical ones are those come from program analysis, in which {\tool} has comparable performance to the best tool.

From Table~\ref{table:str_int_benchmark}, we can see that {\tool} significantly outperforms all other tools. The total number of failed cases is only a half to Z3, which is ranked the 2nd. In fact, most of the failed tests come from the analysis of one Python core library function that converts an IPv6 address to a number. We generated 2028 tests from this function, and among them around 1000 cases  are too difficult for all solvers.  If we exclude results on this function, then {\tool} has only $218$ failed cases in total. This is significantly better than the 2nd place tool Z3, which fails $942$ cases in total.

@@

For further details of the experiment, we have encoded the checkLuhn algorithm introduced in Section~\ref{section:introduction} for cases of 2 to 12 loops (digits). We ran these tests additionally to the experiments above with the timeout set to 120s. The result is summarized in Table~\ref{table:checkLuhn}.
In these tests, {\tool} can solve all problems within 1s while CVC4 only returns a model for cases of 2 to 5 loops and Z3Str3 could not solve any of these problems (either timeout or UNKNOWN). However, Z3 can still solve 7 out of the 11 problems and the problems got timeout are cases for 4, 5, 7, and 9 loops. The behavior of Z3 is not unexpected, since all the problems are satisfiable and sometimes the solver are lucky and went into a branch with a correct model very quickly.

@@

The second benchmark is called APLAS that was created by authors of \textsf{$Kepler_{22}$}~\cite{aplas}. The benchmark includes a total of 600 hand-drafted tests (298 satisfiable and 302 unsatisfiable) involving looping word equations (Both sides of the string equality have a common variable) and length constraints over strings.

>>>>

We have implemented our string constraint solving procedure in a tool called {\tool}. {\tool} is implemented as a theory solver of the SMT solver Z3~\cite{z3}. In this way, we can concentrate on solving conjunctive constraints and let Z3 handle the other boolean connectives. Secondly, it makes it possible to solve not only formulae over string constraints but also combinations of string constraints with other theories that Z3 supports. Furthermore, this approach allows us to more effectively handle the arithmetic constraints that are generated by the under-approximation module and, lastly, it eliminates the need to have our own parser.

In {\tool}, we use the following PFA selection strategy. We use \emph{numeric PFAs} for string variables appearing in string-number conversion and \emph{standard PFAs} for others. We select a size $m$ for numeric PFAs, a number $p$ of their loops, and the length $q$ of the loops. Initially, we set $(m,p,q)=(5,2,q)$ where $q$ is dynamic and obtained from our internal static analysis. We double $m$ and increase $p$ and $q$ by one if refinement is required. We set an upper bound for each parameter and report UNKNOWN if a solution cannot be found within the bound.

Our over-approximation module also uses heuristics to derive the constant value of any side of the constraint $n=\sti{x}$ to refine the over-approximation. For instance, assume we can derive that $n=12$ from some integer constraints. Then we can derive the value of $x$ belongs to the regular language $(0^*12)$. 
%HERE

%more or less standard
The way our theory solver and Z3 interact is almost standard. When Z3 asks our theory solver a string constraint satisfiability problem, our solver tries to prove it is SAT or UNSAT using the procedure discuss in this paper. For under-approximation, whenever a corresponding linear formula is created, we attach the current value of $m$, $p$, $q$ to the formula, and then push it to Z3 core. If our theory solver reports UNKNOWN, Z3 remembers it in a global flag \textsf{incomplete} and either tries another solution branch, or the same solution branch with different value of $m$, $p$, $q$. If Z3 completes the search of all solution branches, it reports UNSAT if the flag \textsf{incomplete} is down, and UNKNOWN otherwise.

@@

We perform two sets of experiments. In the first set of experiments, we compare {\tool} with other tools on existing benchmarks over basic string constraints. Those benchmarks do not involve string-number conversion function. In the second set of experiments, we compare {\tool} with the other tools on new suites focusing on string-number conversion. Our goals of experiments are the following:
\smallskip

\begin{itemize}
	\item {\tool} performs as good as or better than the other tools in solving the  satisfiability problems of basic string constraints.

@@

	\item {\tool} performs significantly better than the other tools in solving the satisfiability problems on string-number conversion benchmark, and this shows  the efficiency of PFA in general and \emph{numeric} PFA in particular.
\end{itemize}
		\smallskip

@@

\begin{itemize}
	\item PyEx~\cite{pyex} comes from running the symbolic executor PyEx over some Python packages.

@@

	\item \changed{ LeetCode comes from running PyEx over a sample code collected from the LeetCode~\cite{LeetCode} website, including functions that check whether a string is a valid IPv4 or IPv6 address, sum up two binary numbers, check whether an input string is an abbreviation of another input string, and convert a sequence of digits to a string according to a given mapping.}

@@

	\item StringFuzz~\cite{blotsky2018stringfuzz}  is generated by the fuzz testing tool of the same name.

@@

In the second experiment, we compare with tools supporting string-number conversion on the benchmarks collected from the symbolic executor \texttt{Py-Conbyte}\footnote{\url{https://github.com/spencerwuwu/py-conbyte}}, which has the supports of string-number conversion. We ran it on several examples collected from the LeetCode platform and from Python core libraries, which involve diverse usages of string-number conversion in Python such as parsing date-time, verifying and restoring IP addresses from strings, etc. We also have examples that encode execution paths of some JavaScript programs (the Luhn algorithm and some array manipulations).

All experiments were executed on a machine with 4-core CPU, 8 GiB RAM. The timeout was set to 10s for each test.
We use the results from {\tool}, CVC4, and Z3 as the reference answer for the validation of the correctness of the results. Occasionally, two of them report inconsistent  answers (one SAT and one UNSAT). To decide which solver is right, we developed a validator. It takes the model $I$ returned from the solver who reported SAT, assigns $I(x)$ to all variables $x$ in the test to obtain a modified test, and re-evaluates the modified test by multiple solvers. If the results from all solvers are consistent, we mark the test SAT or UNSAT according to the results. Otherwise, we manually simplify and inspect the test until we get a conclusive result.

The results of the experiments are summarized in Table~\ref{table:base_benchmark}, Table~\ref{table:str_int_benchmark}, and Table~\ref{table:checkLuhn}. Rows with heading \texttt{SAT}/\texttt{UNSAT} show numbers of solved formulae. Rows with heading \texttt{UNKNOWN}, \texttt{TIMEOUT}, or \texttt{ERROR} indicate the number of instances for which the solver fails to return an answer. \texttt{INCORRECT} shows the number of cases where the tool gives a wrong answer\footnote{A correct answer is decided by mainly CVC4, then the consensus of {\tool} or Z3.}.

@@

From Table~\ref{table:base_benchmark}, we can see that the performance of {\tool} is as good as that of the most competitive tools such as CVC4 and Z3 on basic string constraints. In all of the benchmarks, {\tool} ranked either the 1st or the 2nd on the number of solved (SAT+UNSAT) cases. On the APLAS and StringFuzz benchmarks that are SAT, {\tool} does not perfomr as well as the best performing tool. We however do not consider this crucial because both of these benchmarks are artificial, either hand-crafted or randomly generated. On the most important benchmarks, those that come from program analysis, {\tool} is comparable to the best performing tool.

From Table~\ref{table:str_int_benchmark}, we can see that {\tool} significantly outperforms all other tools. It fails on twice less cases then Z3, which is ranked the 2nd. In fact, most of the failed tests come from the analysis of one Python core library function that converts an IPv6 address to a number. From around 2028 tests generated from this function, around 1000 tests are too difficult for all solvers.  If we exclude the 2028 tests generated from this function, then {\tool} has only $218$ failed cases in total. This is significantly better than the 2nd place tool Z3, which fails on $942$ cases in total.

@@

As an addition experiment, we have encoded the checkLuhn algorithm introduced in Section~\ref{section:introduction} for the cases with 2 to 12 loops (digits). We ran these tests 
%additionally to the experiments above 
with the timeout set to 120s. The result is summarized in Table~\ref{table:checkLuhn}.
In these tests, {\tool} can solve all problems within 1s while CVC4 only returns a model for cases of 2 to 5 loops and Z3Str3 could not solve any of these problems (either timeout or UNKNOWN). However, Z3 can still solve 7 out of the 11 problems, while timeouting in the cases with 4, 5, 7, and 9 loops. The behavior of Z3 is not entirely unexpected. All the problems are satisfiable and the solver may be lucky to guess the solution quickly. % branch with a correct model quickly.

@@

The second benchmark is called APLAS that was created by authors of \textsf{$Kepler_{22}$}~\cite{aplas}. The benchmark includes a total of 600 hand-crafted tests (298 satisfiable and 302 unsatisfiable) involving looping word equations (Both sides of the string equality have a common variable) and length constraints over strings.

### 2020-03-30 final-version/main.tex
little changes in the text

%HERE
The straightforward extension of existing FA based techniques boils down to solving the satisfiability problem of integer polynomial-exponential constraints, whose decidability is still open (Section~\ref{section:s2i}).

@@

\changed{
PASS\cite{goudong2013pass} uses quantified formulae over arrays with symbolic length to encode string constraints, and a specialized quantifier elimination to solve them. Though it differs from our approach significantly and handling the quantification is expensive, PASS is indeed similar in how finite automata and string-number conversion constraints are translated to formulae. But our translation is in general more efficient because it uses only linear arithmetics. In contrast, PASS translates string constraints to formulae with quantifier and array predicates. We did not compare with PASS empirically. The authors informed us that PASS is no longer maintained and is not open-source.
}

@@

\begin{acks}
This research was  partially supported by the  Swedish Foundation for Strategic Research and Ministry of Science and Technology Project,Taiwan (no. 106-2221-E-001 -009 -MY3).
\end{acks}

>>>>

%%%%This sounds like contradicting the next paragraph. I think it is not necessary here, it is not a related work, so I am just commenting out.
%The straightforward extension of existing FA based techniques boils down to solving the satisfiability problem of integer polynomial-exponential constraints, whose decidability is still open (Section~\ref{section:s2i}).

@@

\changed{
PASS \cite{goudong2013pass} uses quantified formulae over arrays with symbolic length to encode string constraints, and a specialized quantifier elimination to solve them. Though it differs from our approach significantly and handling the quantification is expensive, PASS is indeed similar in how finite automata and string-number conversion constraints are translated to formulae. Our translation is in general more efficient because it uses only linear arithmetics. In contrast, PASS translates string constraints to formulae with quantifier and array predicates. We did not compare with PASS empirically. The authors informed us that PASS is no longer maintained and is not open-source.
}

@@

\begin{acks}
This research was  partially supported by the  Swedish Foundation for Strategic Research and Ministry of Science and Technology Project,Taiwan (no. 106-2221-E-001 -009 -MY3), Ministry of Education, Youth and Sports of Czech Republic (no. LL1908).
\end{acks}
